---
title: "Getting started with dastash"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with dastash}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

dastash keeps R values on disk, under keys, for as long as you want them and no
longer. A stash is a directory: open it, read and write it, and every R process on the
machine that opens the same directory sees the same entries.

```{r}
library(dastash)

dir <- file.path(tempdir(), "getting-started")
s <- stash(dir)
s
```

## Reading and writing

`stash_set()` stores a value under a key and returns the stash, so writes chain.
`stash_get()` reads it back.

```{r}
s |>
  stash_set("greeting", "hello") |>
  stash_set("numbers", 1:10)

stash_get(s, "numbers")
```

A key with no entry is an error, unless you supply a `default`. A stored `NULL` is a
value like any other, so the two never get confused.

```{r}
try(stash_get(s, "absent"))
stash_get(s, "absent", default = NA)
stash_has(s, c("greeting", "absent"))
```

`stash_mget()` and `stash_mset()` read and write several entries at once, and
`stash_delete()` removes entries.

```{r}
stash_mset(s, list(a = 1, b = 2))
stash_mget(s, c("a", "b"))
stash_delete(s, c("a", "b"))
```

## Keys

A string is its own key, so a `"group/item"` convention gives you prefixes to list by:

```{r}
stash_set(s, "prices/XSWX", 101.2)
stash_set(s, "prices/XNYS", 57.8)
stash_keys(s, prefix = "prices/")
```

A key can also be a number, a date, a vector or a named list. It is written in a fixed
text encoding: names are sorted, `1L` and `1` agree, and a time zone is not part of a
time's identity, so the same values always find the same entry.

```{r}
stash_set(s, list(exchange = "XSWX", date = as.Date("2026-08-29")), "quotes")
stash_get(s, list(date = as.Date("2026-08-29"), exchange = "XSWX"))
stash_key_chr(list(exchange = "XSWX", date = as.Date("2026-08-29")))
```

## Expiry and tags

An entry can expire: after a number of seconds, a `difftime`, or at a `POSIXct` time. An
expired entry is gone for every reader at once.

```{r}
stash_set(s, "session-token", "abc123", expire = as.difftime(30, units = "mins"))
stash_set(s, "stale", 1, expire = 0)
stash_has(s, "stale")
```

Tags group entries for invalidation. Tag everything derived from one source, and drop
it all when that source changes:

```{r}
stash_set(s, "report/q3", "...", tags = "ledger")
stash_set(s, "report/q4", "...", tags = "ledger")
stash_evict(s, tag = "ledger")
stash_keys(s, prefix = "report/")
```

## Memoising a function

`stash_memoise()` wraps a function so that each call is looked up first and computed
only on a miss. The results outlive the session, and every process using the stash
shares them.

```{r}
slow_mean <- function(n, seed = 1) {
  Sys.sleep(0.5)
  set.seed(seed)
  mean(rnorm(n))
}
fast_mean <- stash_memoise(slow_mean, s)

system.time(fast_mean(1e5))
system.time(fast_mean(1e5))
```

A call is keyed on the function's name and its arguments, with defaults filled in, so
`fast_mean(1e5)` and `fast_mean(1e5, seed = 1)` are the same entry. The key does not
depend on the function's code: when its meaning changes, pass a new `version`.

```{r}
stash_memoise_key(fast_mean, 1e5)
stash_forget(fast_mean, 1e5)
```

## Large values

Values of `inline_max` bytes (32 KiB) or more are stored as read-only files, named by
the hash of their content, so identical values are stored once. `stash_path()` gives
the file, for tools that read it in place.

```{r}
big <- runif(1e5)
stash_set(s, "big", big)
path <- stash_path(s, "big")
file.size(path)
identical(readRDS(path), big)
```

## Looking inside

```{r}
stash_entries(s)[, c("key", "bytes", "inline", "expires")]
stash_stats(s)[, c("count", "bytes_inline", "bytes_blob", "volume")]
```

```{r}
stash_close(s)
unlink(dir, recursive = TRUE)
```
