---
title: "Operating a shared stash"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Operating a shared stash}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

A stash is meant to be shared: by interactive sessions, scheduled jobs and parallel
workers on one machine, all at once. This article covers what that asks of you, and
what dastash guarantees in return.

```{r}
library(dastash)
dir <- file.path(tempdir(), "shared")
s <- stash(dir, size_limit = 1024^2)
```

## Where a stash can live

On a **local filesystem** of the machine whose processes use it. The engine underneath,
'libmdbx', memory-maps the database and coordinates writers through a lock file, and
both need a local filesystem. A stash on a network filesystem (NFS, SMB, a cloud file
share) is not supported.

## Many processes

Any number of processes can read at once, and never wait: a read sees the last
committed state. Writes take turns. A write that waits longer than `timeout` seconds
(60 by default) for another process's write raises `dastash_busy`, not a hang.

Each process opens the stash itself. A handle cannot be carried into a forked child,
such as a `parallel::mclapply()` worker. Using it there raises `dastash_forked`, so open
the stash inside the worker:

```{r, eval = FALSE}
parallel::mclapply(keys, function(k) {
  s <- stash(dir)
  on.exit(stash_close(s))
  stash_get(s, k, default = NULL)
})
```

Workers started as fresh processes ('callr', 'mirai', 'future' with multisession) have
nothing to inherit, and open the stash like any other process.

Every single operation is atomic across processes. Eight processes calling
`stash_incr()` on one key count eight. `stash_add()` writes only if there is no entry.
`stash_pop()` reads and deletes in one step.

```{r}
stash_incr(s, "jobs-run")
stash_add(s, "lock/nightly", Sys.getpid(), expire = 3600)
stash_add(s, "lock/nightly", Sys.getpid())
```

## Transactions

`stash_transact()` makes a block of writes atomic: all commit, or none do.

```{r}
stash_set(s, "balance", 100)
stash_transact(s, {
  stash_set(s, "balance", stash_get(s, "balance") - 30)
  stash_incr(s, "withdrawals")
})
stash_get(s, "balance")
```

It holds the write lock while it runs, so other processes' writes wait for it. Keep it
short, and never compute something slow inside it. It is also the fastest way to write
many entries: making a write durable is the expensive part, and a transaction pays it
once.

Memoised functions follow the same rule. Two processes that miss the same call at the
same moment both compute it, and one result is kept. That is duplicated work, never a
wrong answer.

## Durability

By default (`durability = "safe"`) every write is on disk before it returns.
`durability = "fast"` skips waiting for the disk: a crash of the machine can lose recent
writes, but never damages the stash. `"unsafe"` can damage it, for stashes you can
rebuild. The first process to open a stash sets this for every process that opens it
while it is open; `stash_stats()` reports what is in force.

```{r}
stash_stats(s)$durability
```

## Staying within a size limit

`size_limit` bounds what the entries hold. A write that goes over it evicts the least
recently stored entries, a few at a time, so the limit holds without a background
process. `stash_cull()` finishes the job at once. Expired entries go first.

```{r}
for (i in 1:5) stash_set(s, paste0("frame/", i), runif(5e4))
stash_count(s)
stash_cull(s)
stash_count(s)
```

The database file keeps the pages it frees for reuse rather than shrinking, so
`stash_volume()`, which reports bytes on disk, can stay above the limit.

## Checking and repairing

A process that dies in the middle of a write can leave, at most, a file that nothing
refers to. `stash_check()` finds that, and any other damage. `repair = TRUE` fixes it,
and `hash = TRUE` also verifies every stored file's content.

```{r}
stash_check(s)
stash_check(s, repair = TRUE)
```

Run it now and then from a maintenance job, as you would `stash_cull()`.

```{r}
stash_close(s)
unlink(dir, recursive = TRUE)
```
